MS Regressions

### 2021-12-01 msdepression_regressions_started_20211118.Rmd
figuring out why some people are broken

#goal is to keep people who were seen by 
 data_empi_acc_f_phq <- data %>% 
   filter(Provider %in% ms_providers) %>%       #n=17,067
   filter(grepl("G35", ICD10)) %>%              #n=16830
   mutate(across(.cols = columns_to_make_integer, .fns = as.integer)) %>%
   mutate(sex_binarized = ifelse(SEX == "MALE", 1, 2)) %>%
   mutate(osex = ordered(sex_binarized,levels = c(1,2), labels = c("Male","Female"))) %>%
   mutate(race_binarized = ifelse(RACE == "WHITE", 1, 2)) %>%
   mutate(orace = ordered(race_binarized,levels = c(1,2), labels = c("White","Non-white"))) %>%
   mutate(EXAM_DATE = as.Date(BEGIN_EXAM_DTTM, format = "%m/%d/%y")) %>% 
   mutate(EXAM_DATE = gsub(EXAM_DATE, pattern = "-", replacement = ""))

@@

#fdr corrected
fascicle_agedep_anova_fdr <- fdr_anova_generic(fascicle_agedep_anova, 3)
print(fascicle_agedep_anova_fdr)

@@

>>>>

#goal is to keep people who were seen by 
 data_empi_acc_f_phq <- data %>% 
   filter(Provider %in% ms_providers) %>%       #n=17,067
   filter(grepl("G35", ICD10)) %>%              #n=16830
   mutate(across(.cols = columns_to_make_integer, .fns = as.integer)) %>%
   mutate(sex_binarized = ifelse(SEX == "MALE", 1, 2)) %>%
   mutate(osex = ordered(sex_binarized,levels = c(1,2), labels = c("Male","Female"))) %>%
   mutate(race_binarized = ifelse(RACE == "WHITE", 1, 2)) %>%
   mutate(orace = ordered(race_binarized,levels = c(1,2), labels = c("White","Non-white"))) %>%
   mutate(EXAM_DATE = as.Date(BEGIN_EXAM_DTTM, format = "%m/%d/%y")) %>% 
   mutate(EXAM_DATE = gsub(EXAM_DATE, pattern = "-", replacement = "")) %>%
   mutate(EMPI = as.factor(EMPI))

@@

#fdr corrected - depression alone
fascicle_agedep_anova_fdr <- fdr_anova_generic(fascicle_agedep_anova, 2)
print(fascicle_agedep_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,2)
print(fascicle_anova_fdr_association)

#fdr corrected - interaction
fascicle_agedep_anova_fdr <- fdr_anova_generic(fascicle_agedep_anova, 3)
print(fascicle_agedep_anova_fdr)

@@

```{r age_by_depression_unique}

#lm
fascicle_agedep_lm <- lapply(fascicle_names, function(x) 
{
   lm(substitute(i ~ PAT_AGE_AT_EXAM*depGroupVar, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
# lm(substitute(i ~ depGroupVar + PAT_AGE_AT_EXAM + osex, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
})
names(fascicle_agedep_lm) <- fascicle_names

#fdr corrected - depression alone
fascicle_agedep_anova_fdr <- fdr_anova_generic(fascicle_agedep_anova, 2)
print(fascicle_agedep_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,2)
print(fascicle_anova_fdr_association)

#fdr corrected - interaction
fascicle_agedep_anova_fdr <- fdr_anova_generic(fascicle_agedep_anova, 3)
print(fascicle_agedep_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,3)
print(fascicle_anova_fdr_association)
```
```{r depression_unique_all_fascicles_together}

#isolate out only the depressed group and healthy group, make a column that sums up the total prop lost)
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,]  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names))) %>%
  ungroup()
##

dep_vs_healthy <- lm(total_fascicle_prop_lost ~ depGroupVar + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)

ggplot(data = dep_and_healthy_groups_for_ICD_analysis, aes(x=PAT_AGE_AT_EXAM,y=total_fascicle_prop_lost, color = EMPI, group = EMPI)) + 
  geom_point(position = "jitter", show.legend = F) + 
  geom_line(show.legend = F) + 
  facet_wrap(~EMPI)

#lm
fascicle_agedep_lm <- lapply(fascicle_names, function(x) 
{
   lm(substitute(i ~ PAT_AGE_AT_EXAM*depGroupVar, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
# lm(substitute(i ~ depGroupVar + PAT_AGE_AT_EXAM + osex, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
})
names(fascicle_agedep_lm) <- fascicle_names

#fdr corrected - depression alone
fascicle_agedep_anova_fdr <- fdr_anova_generic(fascicle_agedep_anova, 2)
print(fascicle_agedep_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,2)
print(fascicle_anova_fdr_association)

#fdr corrected - interaction
fascicle_agedep_anova_fdr <- fdr_anova_generic(fascicle_agedep_anova, 3)
print(fascicle_agedep_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,3)
print(fascicle_anova_fdr_association)
```

```{r depression_mm_all_fascicles_together}

#isolate out only the depressed group and healthy group, make a column that sums up the total prop lost)
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles[df_demo_and_fascicles$depGroupVar != 0,]  %>% 
  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names))) %>%
  ungroup()
##

### thank you Bart! 11/30/2021
ggplot(data = dep_and_healthy_groups_for_ICD_analysis, aes(x=PAT_AGE_AT_EXAM,y=total_fascicle_prop_lost, color = EMPI, group = EMPI)) + 
  geom_point(position = "jitter", show.legend = F) + 
  geom_line(show.legend = F) + 
  facet_wrap(~depGroupVar)

ggplot(data = dep_and_healthy_groups_for_ICD_analysis, aes(x=PAT_AGE_AT_EXAM,y=total_fascicle_prop_lost, color = EMPI, group = EMPI)) + 
  geom_point(position = "jitter", show.legend = F) + 
  geom_line(show.legend = F) + 
  facet_wrap(~EMPI)

dep_vs_healthy <- lmerTest::lmer(total_fascicle_prop_lost ~ depGroupVar + (1 | EMPI) + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)

dep_easy <- lm(total_fascicle_prop_lost ~ depGroupVar + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)

ggplot(data = dep_and_healthy_groups_for_ICD_analysis2, aes(x=`(Intercept)`, y=total_fascicle_prop_lost, color = EMPI, group = EMPI)) + 
  geom_point(position = "jitter", show.legend = F) + 
  geom_line(show.legend = F) + 
  facet_wrap(~EMPI)

#lm
fascicle_agedep_lm <- lapply(fascicle_names, function(x) 
{
   lmerTest::lmer(substitute(i ~ depGroupVar + (1|EMPI) + PAT_AGE_AT_EXAM , list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
# lm(substitute(i ~ depGroupVar + PAT_AGE_AT_EXAM + osex, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
})
names(fascicle_agedep_lm) <- fascicle_names

#fdr corrected - depression alone
fascicle_agedep_anova_fdr <- fdr_anova_generic(fascicle_agedep_anova, 2)
print(fascicle_agedep_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,2)
print(fascicle_anova_fdr_association)

#fdr corrected - interaction
fascicle_agedep_anova_fdr <- fdr_anova_generic(fascicle_agedep_anova, 3)
print(fascicle_agedep_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,3)
print(fascicle_anova_fdr_association)
```
